---
title: "Reproducible Research Course Project 1"
author: "Devarshi Mukherji"
date: "7/11/2020"
output: html_document
---

## Read in the data
``` {r, echo = TRUE}
setwd("C:/Devarshi's Work/Coursera/Reproducible Research/Week 2/repdata_data_activity")
data <- read.csv("activity.csv")
```

Exploratory Analysis of the Data
```{r}
names(data)
str(data)
head(data)
sum(is.na(data$steps))
```

## Create Histogram of number of steps in a day
``` {r, echo = TRUE}
stepsAday <- aggregate(data$steps ~ data$date, data, sum, na.rm = TRUE)
png("plot1.png", width = 480, height = 480)
hist(stepsAday$`data$steps`)
dev.off()
hist(stepsAday$`data$steps`)
```

## Caclculate Mean and Median Steps taken per day
``` {r, echo = TRUE}
meanSteps <- mean(stepsAday$`data$steps`)
print(meanSteps)
medianSteps <- median(stepsAday$`data$steps`)
print(medianSteps)
```

## Time series plot of average number of steps taken
``` {r, echo = TRUE}
meanStepsAday <- aggregate(steps ~ interval, data, mean, na.rm = TRUE)
png("plot2.png", width = 480, height = 480)
plot(steps~interval, meanStepsAday, type = "l")
dev.off()
plot(steps~interval, meanStepsAday, type = "l")
```

## Find the 5 minute interval that contains the max number of steps on average
``` {r, echo = TRUE}
maxSteps <- meanStepsAday[which.max(meanStepsAday$steps),]$interval
print(maxSteps)
```

## Imputing missing Data
``` {r, echo = TRUE}
missings <- sum(is.na(data$steps))
print(missings)
averageOfInterval <- function (interval) { 
  meanStepsAday[meanStepsAday$interval==interval, ]$steps  
}
imputedData <- data
for (i in 1:nrow(imputedData)) {
  if(is.na(imputedData[i,]$steps)) {
    imputedData[i,]$steps <- averageOfInterval(imputedData[i,]$interval)
  }
}
```

## Histogram of the total number of steps taken each day after missing values are imputed
``` {r, echo = TRUE}
imputedStepsAday <- aggregate(steps ~ date, imputedData, sum)
png("plot3.png", width = 480, height = 480)
hist(imputedStepsAday$steps)
dev.off()
hist(imputedStepsAday$steps)
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
``` {r, echo  = TRUE}
imputedData$date <- as.Date(strptime(imputedData$date, format = "%Y-%m-%d"))
imputedData$day <- weekdays(imputedData$date)
for (i in 1:nrow(imputedData)) {
  if (imputedData[i,]$day %in% c("Saturday", "Sunday")) {
    imputedData[i,]$day <- "weekend"
  } else {
    imputedData[i,]$day <- "weekday"
  }
}
weekSplit <- aggregate(imputedData$steps ~ imputedData$interval + imputedData$day, imputedData, mean)
names(weekSplit) <- c("interval", "day", "steps")
png("plot4.png", width = 480, height = 480)
library(lattice)
xyplot(steps ~ interval | day, weekSplit, type = "h", layout = c(1, 2), xlab = "time", ylab = "steps")
dev.off()
xyplot(steps ~ interval | day, weekSplit, type = "h", layout = c(1, 2), xlab = "time", ylab = "steps")
```